### 2009-10-27 doc/article.tex
added structure for documentation

\documentclass[german]{article}

\usepackage[utf8]{inputenc}
\usepackage[ngerman]{babel}

\usepackage{listings}
\usepackage{color}

\author{Hong-Khoan Quach, Marek Kubica, Julius Adorf \\ Technische Universität München}

\title{ETI-Großprojekt 8\\
       Sommersemester 2009 \\
       {\bf Das Carrierpigeon-Projekt}
}

\date{26.10.2009}

% cheat a little bit
% code samples should not be split
\setlength{\headheight}{-20pt}
\setlength{\textheight}{590pt}

% colorize code samples
\lstset{
    basicstyle=\small,
    keywordstyle=\color[rgb]{0.1,0.1,0.8}
    }

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Motivation}

\section{Projektziel}
\section{Architektur}
\section{Herausforderungen}
\section{Toolchain}
\section{Fazit}

\newpage

\appendix

\newpage

% include sources that are not cited explicitly
% \nocite{}

\bibliographystyle{plain}
\bibliography{article}

\end{document}